---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dpi = 150
)
```
# isoband

[![Build Status](https://travis-ci.org/clauswilke/isoband.svg?branch=master)](https://travis-ci.org/clauswilke/isoband)
[![Coverage Status](https://img.shields.io/codecov/c/github/clauswilke/isoband/master.svg)](https://codecov.io/github/clauswilke/isoband?branch=master)


Generate contour lines (isolines) and contour polygons (isobands) from regularly spaced grids containing elevation data.
  
## Installation

Install from github with:

``` r
devtools::install_github("clauswilke/isoband")
```

## Examples

The two main workhorses of the package are the functions `isolines()` and `isobands()`, respectively. They return a list of isolines/isobands for each isolevel specified. Each isoline/isoband consists of vectors of x and y coordinates, as well as a vector of ids specifying which sets of coordinates should be connected. This format can be handed directly to `grid.polyline()`/`grid.path()` for drawing. However, we can also convert the output to spatial features and draw with ggplot2 (see below).

```{r basic-example, fig.asp = 1}
library(isoband)

m <- matrix(c(0, 0, 0, 0, 0,
              0, 1, 2, 1, 0,
              0, 1, 2, 0, 0,
              0, 1, 0, 1, 0,
              0, 0, 0, 0, 0), 5, 5, byrow = TRUE)

isolines(1:ncol(m), 1:nrow(m), m, 0.5)

isobands(1:ncol(m), 1:nrow(m), m, 0.5, 1.5)
```

The function `plot_iso()` is a convenience function for debugging and testing.
```{r basic-example-plot, fig.asp = 1, out.width = "50%"}
plot_iso(m, 0.5, 1.5)
```

A few more simple examples. Missing values and disconnected areas are all handled correctly.
```{r basic-plotting, fig.asp = 1, out.width = "50%"}
m <- matrix(c(0, 0, 0, 0, 0, 0,
              0, 0, 0, 1, 1, 0,
              0, 0, 1, 1, 1, 0,
              0, 1, 1, 0, 0, 0,
              0, 0, 0, 1, 0, 0,
              0, 0, 0, 0, 0, 0), 6, 6, byrow = TRUE)
plot_iso(m, 0.5, 1.5)

m <- matrix(c(NA, NA, NA, 0, 0, 0,
              NA, NA, NA, 1, 1, 0,
               0,  0,  1, 1, 1, 0,
               0,  1,  1, 0, 0, 0,
               0,  0,  0, 1, 0, 0,
               0,  0,  0, 0, 0, 0), 6, 6, byrow = TRUE)
plot_iso(m, 0.5, 1.5)

m <- matrix(c(0, 0, 1, 1,
              0, 1, 1, 1,
              1, 1, 0, 0,
              0, 0, 0.8, 0), 4, 4, byrow = TRUE)
plot_iso(m, 0.5, 1.5)
```

The algorithm has no problem with larger datasets. Let's calculate isolines and isobands for the volcano dataset, convert to sf, and plot with ggplot2.
```{r volcano, fig.asp = 1, out.width = "75%"}
library(ggplot2)
library(sf)

m <- volcano
b <- isobands((1:ncol(m))/(ncol(m)+1), (nrow(m):1)/(nrow(m)+1), m, 10*(9:19), 10*(10:20))
l <- isolines((1:ncol(m))/(ncol(m)+1), (nrow(m):1)/(nrow(m)+1), m, 10*(10:19))

bands <- iso_to_sfg(b)
data_bands <- st_sf(
  level = 1:length(bands),
  geometry = st_sfc(bands)
)
lines <- iso_to_sfg(l)
data_lines <- st_sf(
  level = 2:(length(lines)+1),
  geometry = st_sfc(lines)
)

ggplot() +
  geom_sf(data = data_bands, aes(fill = level), color = NA, alpha = 0.7) +
  geom_sf(data = data_lines, color = "black") +
  scale_fill_viridis_c(guide = "none") +
  coord_sf(expand = FALSE)

```

We can also use this approach to convert bitmap images into polygons and plot with ggplot2.

```{r polygon-hadley, fig.asp = 1}
library(magick)
library(sf)
library(ggplot2)
library(isoband)
library(colorspace)

sf_from_image <- function(image) {
  image_gray <- image %>% image_quantize(colorspace = "gray")
  image_raster <- as.raster(image_gray)
  d <- dim(image_raster)
  m <- matrix(c((255-col2rgb(image_raster)[1,])), nrow = d[1], ncol = d[2], byrow = TRUE)
  b <- isobands(1:d[2], d[1]:1, m, 20*(-1:13), 20*(0:14))
  bands <- iso_to_sfg(b)
  data <- st_sf(
    level = letters[1:length(bands)],
    geometry = st_sfc(bands)
  )
}

img <- image_resize(image_read("https://pbs.twimg.com/profile_images/905186381995147264/7zKAG5sY_400x400.jpg"), "200x200")
img_sf <- sf_from_image(img)

p1 <- ggplot(img_sf) + 
  geom_sf(color = "gray10", fill = NA, size = 0.05) + 
  coord_sf(expand = FALSE) +
  theme_gray() +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks.length = grid::unit(0, "pt"),
    plot.margin = margin(0, 0, 0, 0)
  )

p2 <- ggplot(img_sf) + 
  geom_sf(aes(fill = level, color = level)) + 
  coord_sf(expand = FALSE) +
  theme_void()

p3 <- ggplot(img_sf) + 
  geom_sf(aes(fill = level), color = "gray30", size = 0.1) + 
  scale_fill_hue(aesthetics = c("color", "fill"), guide = "none", direction = -1) +
  coord_sf(expand = FALSE) +
  theme_void()

cowplot::plot_grid(
  p1,
  p2 + scale_fill_discrete_sequential(aesthetics = c("color", "fill"), palette = "Inferno", guide = "none", rev = TRUE),
  p2 + scale_fill_discrete_sequential(aesthetics = c("color", "fill"), palette = "Viridis", guide = "none", rev = TRUE),
  p3,
  scale = 0.9
)
```

## Performance

The code is written in C++ and performance is generally good. Isolining is about as fast as `grDevices::contourLines()`, isobanding is approximately 2.5 times slower.

```{r benchmark}
microbenchmark::microbenchmark(
  grDevices::contourLines(1:ncol(volcano), 1:nrow(volcano), volcano, levels = 10*(10:18)),
  isolines(1:ncol(volcano), 1:nrow(volcano), volcano, 10*(10:18)),
  isobands(1:ncol(volcano), 1:nrow(volcano), volcano, 10*(9:17), 10*(10:18))
)

```
