---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "50%"
)
```
# isoband

[![Build Status](https://travis-ci.org/clauswilke/isoband.svg?branch=master)](https://travis-ci.org/clauswilke/isoband)
[![Coverage Status](https://img.shields.io/codecov/c/github/clauswilke/isoband/master.svg)](https://codecov.io/github/clauswilke/isoband?branch=master)


Generate contour lines (isolines) and contour polygons (isobands) from regularly spaced grids containing elevation data.
  
## Installation

Install from github with:

``` r
devtools::install_github("clauswilke/isoband")
```

## Example

We begin with a few simple examples. The two main workhorses of the package are the functions `isolines()` and `isobands()`, respectively. They return a list of isolines/isobands for each isolevel specified. Each isoline/isoband consists of vectors of x and y coordinates, as well as a vector of ids specifying which sets of coordinates should be connected. This format can be handed directly to `grid.polyline()`/`grid.path()` for drawing (see below).

```{r basic-example, fig.asp = 1}
library(isoband)

m <- matrix(c(0, 0, 0, 0, 0,
              0, 1, 2, 1, 0,
              0, 1, 2, 0, 0,
              0, 1, 0, 1, 0,
              0, 0, 0, 0, 0), 5, 5, byrow = TRUE)

isolines(1:ncol(m), 1:nrow(m), m, 0.5)
isobands(1:ncol(m), 1:nrow(m), m, 0.5, 1.5)
```

The function `plot_iso()` is a convenience function for debugging and testing.
```{r basic-example-plot, fig.asp = 1}
plot_iso(m, 0.5, 1.5)
```

A few more simple examples. Missing values and disconnected areas are all handled correctly.
```{r basic-plotting, fig.asp = 1}
m <- matrix(c(0, 0, 0, 0, 0, 0,
              0, 0, 0, 1, 1, 0,
              0, 0, 1, 1, 1, 0,
              0, 1, 1, 0, 0, 0,
              0, 0, 0, 1, 0, 0,
              0, 0, 0, 0, 0, 0), 6, 6, byrow = TRUE)
plot_iso(m, 0.5, 1.5)

m <- matrix(c(NA, NA, NA, 0, 0, 0,
              NA, NA, NA, 1, 1, 0,
               0,  0,  1, 1, 1, 0,
               0,  1,  1, 0, 0, 0,
               0,  0,  0, 1, 0, 0,
               0,  0,  0, 0, 0, 0), 6, 6, byrow = TRUE)
plot_iso(m, 0.5, 1.5)

m <- matrix(c(0, 0, 1, 1,
              0, 1, 1, 1,
              1, 1, 0, 0,
              0, 0, 0.8, 0), 4, 4, byrow = TRUE)
plot_iso(m, 0.5, 1.5)
```

The algorithm has no problem with larger datasets. Next we calculate a few isolines/isobands for the volcano dataset.
```{r volcano, fig.asp = 1, out.width = "75%"}
m <- volcano
b <- isobands((1:ncol(m))/(ncol(m)+1), (nrow(m):1)/(nrow(m)+1), m, c(80, 120, 150), c(110, 140, 152))
l <- isolines((1:ncol(m))/(ncol(m)+1), (nrow(m):1)/(nrow(m)+1), m, c(110, 120, 140, 150, 152))

library(grid)
grid.newpage()
grid.path(b[[1]]$x, b[[1]]$y, b[[1]]$id, gp = gpar(fill = "cornsilk", col = NA))
grid.path(b[[2]]$x, b[[2]]$y, b[[2]]$id, gp = gpar(fill = "lightblue", col = NA))
grid.path(b[[3]]$x, b[[3]]$y, b[[3]]$id, gp = gpar(fill = "tomato", col = NA))
for (i in seq_along(l)) {
  grid.polyline(l[[i]]$x, l[[i]]$y, l[[i]]$id)
}
```

Isolining is about as fast as `grDevices::contourLines()`, isobanding is about 2.5 times slower.

```{r benchmark}
microbenchmark::microbenchmark(
  grDevices::contourLines(1:ncol(volcano), 1:nrow(volcano), volcano, levels = 10*(10:18)),
  isolines(1:ncol(volcano), 1:nrow(volcano), volcano, 10*(10:18)),
  isobands(1:ncol(volcano), 1:nrow(volcano), volcano, 10*(9:17), 10*(10:18))
)

```

